---
title: "Lab 02 - Data Wrangling with `dplyr`"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

# Do not adjust these settings
# as they are here to set up 
# a few things for you.

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

```{r load_packages}
# make sure you have installed the DWVpack package
# be sure to review the Getting Started with R chapter 
# if you have questions about how to install DWVpack

# also, make sure you have installed dplyr
# install.packages( "dplyr" )

# Note that you cannot knit with an install.packages() function
# called in your RMD file. You only have to install packages
# once, then they are loaded using the library() function.

library( dplyr )
library( DWVpack )

```

<br>

# Setup

```{r clear_workspace}

# clear the workspace by removing any existing objects
rm( list = ls() )

# you can also restart your session as well

```

<br>

# Part 1: Victims of Murder

## 1. Identify the number of observations, number of variables, and what a single row represents.

```{r}

# To get a sense of the data, use the View() function
View( troubles_fatalities )

# Now add the rest of the code here

```

## 2. Use `select()` to create a smaller version of the dataset containing only the `age` and `sex` variables.

```{r}

# Add your code here

```

## 3. Use `distinct()` to investigate the values of `sex` and `affiliation`.

```{r}

# Add your code here

```

## 4. Use `filter()` to a subset of the data that is just **Catholic** victims, then a subset of victims that are **male** *and* **Protestant**, and finally a subset of victims that are **male** *or* **Younger than 18**.

```{r}

# Add your code here

```

## 5. Use `arrange()` to sort the dataset by `age`.

```{r}

# Add your code here

```


## 6. Use `mutate()` to create a new variable that captures whether a victim is a minor (i.e. younger than 18) or an adult (i.e. 18 years or older).

```{r}

# Add your code here

```

## 7. Use `summarize()` to calculate the number of observations (`n()`) in the data and the **mean** for `age`.

```{r}

# Add your code here

```


## 8. Use `group_by()` with `summarize()` to find the **mean** age for victims for each religious affiliation (i.e. use the `affiliation` variable).

```{r}

# Add your code here

```

<br>

----

# Part 2

Now repeat the processes from Part 1 using the data you found in [Lab 1](), substituting the variables in your data. 

## 1. Identify the number of observations, number of variables, and what a single row represents.

```{r}

# To get a sense of the data, use the View() function
View( troubles_fatalities )

# Now add the rest of the code here

```

## 2. Use `select()` to create a smaller version of the dataset...

```{r}

# Add your code here

```

## 3. Use `distinct()` to investigate the values of...

```{r}

# Add your code here

```

## 4. Use `filter()` to a subset of the data that is just **Catholic** victims, then a subset of victims that are **Male** *and* **Protestant**, and finally a subset of victims that are **Male** *or* **Younger than 18**.

```{r}

# Add your code here

```

## 5. Use `arrange()` to sort the dataset by `age`.

```{r}

# Add your code here

```


## 6. Use `mutate()` to create a new variable that captures whether a victim is a minor (i.e. younger than 18) or an adult (i.e. 18 years or older).

```{r}

# Add your code here

```

## 7. Use `summarize()` to calculate the number of observations (`n()`) in the data and the **mean** for `age`.

```{r}

# Add your code here

```


## 8. Use `group_by()` with `summarize()` to find the **mean** age for victims for each religious affiliation (i.e. use the `affiliation` variable).

```{r}

# Add your code here

```
