---
title: "Lab 05 - Bipartite Graphs and Two-Mode Networks"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

# Do not adjust these settings
# as they are here to set up 
# a few things for you.

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>


```{r load_packages}

# load the packages you need
library( SNACpack )
library( network )
library( sna )

```

<br>

# Getting Started

## Setup

```{r}

# clear the workspace
rm( list = ls() )

```

<br>

## Part I: A Conspiracy Network

<br>

### 1. Plot the network using the `gplot()` function.

```{r}

# Your code here

```

<br>

### 2. Based on the plot, what do you observe about the structure of the network? Can you tell which nodes are people and which are locations? What would be your overall assessment of how the network is organized?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 3. Calculate the density of the network.

```{r}

# Your code here

```

<br>

### 4. Interpret the density of the network. What does this number represent in the context of network where individuals are connected to each other through shared places? 

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 5. Is this a "dense" network in your opinion? Why or why not?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 6. Calculate the degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

### 7. Calculate the mean degree centrality scores. Interpret the values.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 8. What do the means suggest about typical levels of connectivity?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 9. Which set of nodes (people or places) has more variation in degree? Why do you think this is the case?   

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 10. Calculate the standardized degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

### 11. Calculate the mean standardized degree centrality score.  

```{r}

# Your code here

```

<br>

### 12. Plot the network using the standardized degree centrality scores. 

```{r}

# Your code here

```

<br>

### 13. Does the visualization change your interpretation of the prompts above? What do you observe that is different (or the same) about the structure of the network given the addition of centrality scores to the plot?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 14. Using the vertex attribute `people_names`, identify which individual is most central in the network.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 15. What might their position in the network suggest about their role in this historical context?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 16. Using the vertex attribute `place_names`, identify which location is most central in the network.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 17. Why might this place have been especially important in terms of connectivity?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 18. Suppose you are an analyst for the British in the 18th century. You are given this network. What is your approach to addressing a potential rebellion in the colonies of British America? Explain.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## Part II: A Kidnapping Network

<br>

### 1. Plot the network using the `gplot()` function.

```{r}

# Your code here

```

<br>

### 2. Based on the plot, what initial patterns do you notice? Are some people involved in many incidents? Are some incidents connected to many people?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 3. Calculate the density of the network.

```{r}

# Your code here

```

<br>

### 4. In this context, what does density tell us? Think in terms of how commonly kidnappers share incidents. Would you describe this as a dense or sparse network? Why?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 5. Is this a "dense" network in your opinion? Why or why not?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 6. Calculate the degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

### 7. Calculate the mean degree centrality scores. Interpret the values.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 8. What does the mean suggest about typical levels of involvement (for people) or complexity (for events)?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 9. Calculate the standardized degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

### 10. Calculate the mean standardized degree centrality score.  

```{r}

# Your code here

```

<br>

### 11. Plot the network using the standardized degree centrality scores. 

```{r}

# Your code here

```

<br>

### 12. What stands out now that didn’t before? Do any individuals or events appear more structurally important? Has your interpretation of the network’s structure changed?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 13. Using the vertex attribute `vertex.names`, identify which individual is most central in the network.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 14. Envision yourself as an analyst trying to reduce kidnapping events (such as these). Based on this network, what is your strategy? Explain.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## Part III: Your Network

For this part of the lab, I would like you to construct a bipartite network. Think about a type of relationship that is interesting, important, etc. and what it connects. It can be a real example from your life, something fictitious, anything really. The *only* restriction is that it should have <u>at least</u> 25 nodes total (e.g., you could have 15 nodes in the first mode and 10 nodes in the second mode). We will be using this network in the next lab, so please be thoughtful in putting your network together.

After creating your bipartite network, do the following:

### 1. Create an adjacency matrix of the network using the `matrix()` function.

```{r}

# Your code here

```

<br>

### 2. Create an object of class `network`. 

```{r}

# Your code here

```

<br>

### 3. Plot the network using the `gplot()` function.

```{r}

# Your code here

```

<br>

### 4. Calculate the density of the network.

```{r}

# Your code here

```

<br>

### 5. Calculate the degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

### 6. Calculate the mean degree centrality scores.

```{r}

# Your code here

```

<br>

### 7. Calculate the mean standardized degree centrality score.  

```{r}

# Your code here

```

<br>

### 8. Plot the network using the standardized degree centrality scores. 

```{r}

# Your code here

```

<br>

### 9. Based on the plot, what would you say about the structure of your network?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## *Challenge Question*

Feeling bold? Try this challenge question. It is not required, but try taking a crack at it.

*  Here is an edgelist for a small 8-node two-mode network:  
    * Percy - C1
    * Percy - C2
    * Percy - C3
    * Harry - C1
    * Harry - C2
    * Isildur - C1
    * Isildur - C2
    * Parker - C2
    * Parker - C3
    * Igor - C3

**Task:** Without using a computer, **or** sketching this network, tell me which two nodes have the highest degree centrality in this network?  

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  
