---
title: "Lab 03 - Tidy Data and Data Transformation"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

# Do not adjust these settings
# as they are here to set up 
# a few things for you.

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

```{r load_packages}

library( DWVpack )
library( dplyr )
library( tidyverse )

```

<br>

# Setup

```{r clear_workspace}

# clear the workspace by removing any existing objects
rm( list = ls() )

# you can also restart your session as well

```

<br>

# Part 1: Tidying Hate Crime Data

## 1. Create a new object called `dat` that contains the `tempe_hate` data.

```{r}

# Add your code here

```

## 2. Determine the number of observations and variables in `dat`, examine the variable names, and identify what a single row in the dataset represents.

```{r}

# Add your code here

```

## 3. Using `select()`, create a new version of `dat` called `dat_tidy` that contains only the following variables (as you select these variables, rename them using the names in parentheses):  

  - incident identifier (`incident_id`)
  - date and time of the crime (`time_date`)
  - type of bias (`bias_type`)
  - type of offense (`offense_type`)

```{r}

# Add your code here

```

## 4. Examine the resulting data and confirm that `dat_tidy` now contains only these variables.

```{r}

# Add your code here

```

## 5. Examine the `time_date` variable. What information is contained in this variable, and how are the values formatted? Use an appropriate function from `lubridate` to parse `time_date` and create a new variable called `datetime`.

```{r}

# Add your code here

```

## 6. Using the `datetime` variable you created in the prior step, create the following new variables (i.e. columns):  

  - `date`, containing only the date
  - `time`, containing the time in hour-and-minute format
  - `year`, containing the year
  - `month`, containing the abbreviated month label
  - `day`, containing the day of the month

```{r}

# Add your code here

```

## 7. The values in `offense_type` contain several different offense descriptions. Some of these descriptions represent similar types of offenses that we would like to place into broader categories. Examine the unique values of `offense_type`.

```{r}

# Add your code here

```

## 8. Based on those values, identify which offenses could reasonably be grouped into the following four categories:  

  - "Assault"
  - "Property Damage"
  - "Theft"
  - "Harassment"

```{r}

# Add your code here

```

## 9. Use `mutate()` and `case_when()` to create a new variable called `offense_type_clean` that groups the original offense descriptions into the four categories from the prior step.

```{r}

# Add your code here

```

## 10. Examine the unique values of both `offense_type` and `offense_type_clean`. How did the transformation change the number and meaning of the offense categories?

```{r}

# Add your code here

```

## 11. Identify the unique values of `bias_type`.

```{r}

# Add your code here

```

## 12. Use `mutate()` and `case_when()` to create a variable called `bias_type_clean` that:  

  - identifies blank values as `Unknown`
  - combines the different categories referring to anti-gay, anti-lesbian, or mixed LGBTQ groups into `Anti-Gay`
  - retains the original value of `bias_type` when none of these conditions are met

```{r}

# Add your code here

```

## 13. Examine the unique values of `bias_type_clean` and compare them with the original `bias_type`.

```{r}

# Add your code here

```

The `DWVpack` package contains a completed version of these transformations called `tidy_tempe_hate`. You will now use this dataset to check your work.

## 14. Compare the dimensions of your `dat_tidy` object with `tidy_tempe_hate`. Do they contain the same number of observations and variables?

```{r}

# Add your code here

```

## 15. Compare the unique values of `offense_type_clean` and the unique values of `bias_type_clean` in the two datasets.

If you find differences between your data and `tidy_tempe_hate`, return to your transformation code and determine what caused the differences.

```{r}

# Add your code here

```

At this point, the code you wrote throughout the lab represents a **data-cleaning recipe**: a reproducible series of steps that transforms the original `tempe_hate` data into an analysis-ready dataset.

## 16. Organize the code from the previous parts into a single workflow that begins with `tempe_hate` and produces your final tidy object called: `dat_tidy`. Run the complete script from beginning to end to confirm that it successfully recreates your cleaned dataset.

```{r}

# Add your code here

```

<br>

----

# Part 2: Tidy Your Own Data

Now that you have worked through the process of cleaning and transforming the `tempe_hate` data, you will apply the same ideas to **your own dataset**. 
The specific transformations you need will depend on the structure and content of your data.

## 1. Identify **at least three variables** in your dataset that could benefit from cleaning or transformation. Explain what, if anything, needs to be changed about each variable.

*ADD YOUR TEXT HERE*

## 2. Create a new object for your tidy dataset. Do not overwrite your original data.

```{r}

# Add your code here

```

## 3. Apply appropriate transformations to your data. Your transformations should use techniques covered in the readings and may include:  

  - selecting and renaming variables
  - cleaning or standardizing character variables
  - combining similar categories
  - creating new variables using logical conditions
  - parsing and transforming dates or times
  - reshaping data when necessary

```{r}

# Add your code here

```

## 4. Organize your transformations into a reproducible **data-cleaning recipe** that starts with your original dataset and produces your final tidy dataset. Run your code from beginning to end to make sure that it works correctly.

```{r}

# Add your code here

```
