---
title: "Lab 06 - Projection of Bipartite Graphs"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

# Do not adjust these settings
# as they are here to set up 
# a few things for you.

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

```{r load_packages}

# load the packages you need
library( SNACpack )
library( network )
library( sna )

```

<br>

# Getting Started

## Setup

```{r}

# clear the workspace
rm( list = ls() )

```

<br>

## Part I: A Conspiracy Network

### 1. Create an object of class `matrix` from the `paul_revere_net` network. Use the `as.sociomatrix()` function in the `network` package to do so. ( **NOTE: DO NOT PRINT OUT THE MATRIX**).

```{r}

# Your code here

```

<br>

### 2. Create the unipartite graph that represents the *person* matrix.

```{r}

# Your code here

```

<br>

### 3. Create the unipartite graph that represents the *group* matrix.

```{r}

# Your code here

```

<br>

### 4. Plot all three networks. Use the `edge_rescale()` and `edge_shade()` functions in the `SNACpack` package to use the edge weights in the plots for each projected graph. How does visualizing edge weight help interpret the structure?

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 5. Create an object of class `network` using the `as.network()` function. Pay attention to the correct usage of the `directed=` argument.

```{r}

# Your code here

```

<br>

### 6. Think about the networked created by the *person* matrix. What do the ties in this network represent? Explain.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 7. Without running any code, describe what kind of actor in this network would have a *high* **degree** centrality score. 

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 8. Describe what kind of actor in this network is likely to have a *low* **closeness** centrality score. 

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 9. Describe the position of a person with a *high* **betweenness** centrality score, but a *low* **degree** centrality score. 

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 10. Calculate the degree, closeness, and betweenness centrality scores for each actor. Feel free to use the `degree()`, `closeness()`, and `betweenness()` functions from the `sna` package.

```{r}

# Your code here

```

<br>

### 11. Calculate the mean degree, closeness, and betweenness centrality scores. Interpret each mean centrality score. What do these means suggest about the structure of the network?

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 12. If we were examining degree centrality, what would a completely centralized network look like? What would its degree centrality scores be? Describe it (or, if you are feeling spirited, create one and plot it!).

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 13. If we were examining betweenness centrality, what would a completely decentralized network look like? What would its betweenness centrality scores be? Describe it (or, if you are still feeling spirited, create one and plot it!).

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 14. Calculate the graph centralization scores for degree, closeness, and betweenness centrality.  

```{r}

# Your code here

```

<br>

### 15. Interpret each graph centralization score. What does each tell us about the distribution of centrality scores in the network? Would you describe the network as centralized or decentralized? Justify your answer using the score.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 16. Plot the network three times using the `gplot()` function where each plot sizes the nodes by the centrality score (i.e. a plot for degree centrality, a plot for closeness centrality, and a plot for betweenness centrality). Examine the plots. Describe how the differences you noted between the centrality measures above are seen in the plots.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 17. Suppose you are an analyst for the British in the 18th century. After analyzing this network, what might you do to address a potential rebellion? How is this informed by your knowledge of the network? Is your answer different based on whether you are examining degree centrality, closeness centrality, or betweenness centrality?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## Part II: A Kidnapping Network

### 1. Create an object of class `matrix` from the `philippine_kidnappings_net` network. Use the `as.sociomatrix()` function in the `network` package to do so. ( **NOTE: DO NOT PRINT OUT THE MATRIX**).

```{r}

# Your code here

```

<br>

### 2. Create the unipartite graph that represents the *person* matrix.

```{r}

# Your code here

```

<br>

### 3. Create the unipartite graph that represents the *group* matrix.

```{r}

# Your code here

```

<br>

### 4. Plot all three networks. Use the `edge_rescale()` and `edge_shade()` functions in the `SNACpack` package to use the edge weights in the plots for each projected graph. How does visualizing edge weight help interpret the structure?

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 5. Create an object of class `network` using the `as.network()` function. Pay attention to the correct usage of the `directed=` argument.

```{r}

# Your code here

```

<br>

### 6. Think about the networked created by the *person* matrix. What do the ties in this network represent? Explain.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 7. Without running any code, describe what kind of actor in this network would have a *high* **degree** centrality score. 

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 8. Describe what kind of actor in this network is likely to have a *low* **closeness** centrality score. 

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 9. Describe the position of a person with a *high* **betweenness** centrality score, but a *low* **degree** centrality score. 

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 10. Calculate the degree, closeness, and betweenness centrality scores for each actor. Feel free to use the `degree()`, `closeness()`, and `betweenness()` functions from the `sna` package.

```{r}

# Your code here

```

<br>

### 11. Calculate the mean degree, closeness, and betweenness centrality scores. Interpret each mean centrality score. What do these means suggest about the structure of the network?

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 12. If we were examining degree centrality, what would a completely centralized network look like? What would its degree centrality scores be? Describe it (or, if you are feeling spirited, create one and plot it!).

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 13. If we were examining betweenness centrality, what would a completely decentralized network look like? What would its betweenness centrality scores be? Describe it (or, if you are still feeling spirited, create one and plot it!).

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 14. Calculate the graph centralization scores for degree, closeness, and betweenness centrality.  

```{r}

# Your code here

```

<br>

### 15. Interpret each graph centralization score. What does each tell us about the distribution of centrality scores in the network? Would you describe the network as centralized or decentralized? Justify your answer using the score.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 16. Plot the network three times using the `gplot()` function where each plot sizes the nodes by the centrality score (i.e. a plot for degree centrality, a plot for closeness centrality, and a plot for betweenness centrality). Examine the plots. Describe how the differences you noted between the centrality measures above are seen in the plots.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

17. Envision yourself as an analyst trying to reduce kidnapping events (such as these). Based on this network, what is your strategy? How is this informed by your knowledge of the network? Is your answer different based on whether you are examining degree centrality, closeness centrality, or betweenness centrality? Explain your answer.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## Part III: Your Network

For this part of the lab, I would like you use the bipartite network you created in [Lab 5](../labs/lab-05-instructions.html). 

For the bipartite network you created, repeat steps 1-4 above.
For the *person* matrix you created in step 2, repeat steps 5-16.
Then answer: 

17. Based on the plot, what would you say about the structure of your network? Is your network centralized or decentralized? Are certain nodes clearly more connected? Any surprising patterns?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## *Challenge Question*

Feeling bold? Try this challenge question. It is not required, but try taking a crack at it.

*  Here is an edgelist for a small 8-node two-mode network:  
    * Percy - C1
    * Percy - C2
    * Percy - C3
    * Harry - C1
    * Harry - C2
    * Isildur - C1
    * Isildur - C2
    * Parker - C2
    * Parker - C3
    * Igor - C3

**Task:** Without using a computer, sketch the bipartite network and create *person* projection. 

Then, answer:

  - Which node(s) have the largest degree centrality?  
  - Which node(s) have the smallest degree centrality?  

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  