---
title: "Lab 04 - Creating Visualizations with `ggplot2`"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

# Do not adjust these settings
# as they are here to set up 
# a few things for you.

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

```{r load_packages}

library( DWVpack )
library( dplyr )
library( tidyverse )
library( ggplot2 )

```

<br>

# Setup

```{r clear_workspace}

# clear the workspace by removing any existing objects
rm( list = ls() )

# you can also restart your session as well

```

<br>

# Part 1: Tidying Bicycle Data

## 1. Create a new object called `dat` that contains the `tucson_bike` data.

```{r}

# Add your code here

```

## 2. Using `select()`, create a new version of `dat` called `dat_tidy` that contains only the following variables:  

  - `accident_id`
  - `time_date`
  - `fatal`

```{r}

# Add your code here

```

## 3. Examine the resulting data and confirm that `dat_tidy` now contains only these variables.

```{r}

# Add your code here

```

## 4. Examine the `time_date` variable. What information is contained in this variable, and how are the values formatted? Use an appropriate function from `lubridate` to parse `time_date` and create a new variable called `datetime`.

```{r}

# Add your code here

```

## 5. Using the `datetime` variable you created in the prior step, create the following new variables (i.e. columns):  

  - `date`, containing only the date
  - `time`, containing the time in hour-and-minute format
  - `year`, containing the year
  - `month`, containing the abbreviated month label
  - `day`, containing the day of the month

```{r}

# Add your code here

```

6. Using your tidied data, create a summary dataset called `bike_monthly` that counts the number of bicycle accidents by year and month.

```{r}

# Add your code here

```

7. Use `ggplot2` to create a visualization of bicycle accidents by month and year. Your plot should:  

  - show the number of bicycle accidents by month
  - separate the data by year
  - use points and lines
  - use color to distinguish years
  - include a title, subtitle, axis labels, and legend label

```{r}

# Add your code here

```

At this point, the code you wrote throughout the lab represents a **data-cleaning and visualization recipe**: a reproducible series of steps that transforms the original `tucson_bike` data into an analysis-ready dataset and visualization.

8. Organize the code from the previous parts into a single workflow that begins with `tucson_bike` and produces your visualization. 
Run the complete script from beginning to end to confirm that it successfully recreates your visualization.

```{r}

# Add your code here

```

<br>

----

# Part 2: Visualizing Your Own Data

Now that you have worked through the process of cleaning and visualizing the `tucson_bike` data, you will apply the same ideas to **your own dataset**. 
The specific visualization you create will depend on the structure and content of your data.

## 1. Identify a question about your dataset that could be explored with a visualization. Your question should focus on a pattern, trend, comparison, or relationship in the data.

*ADD YOUR TEXT HERE*

## 2. Create a new object for the data you will use in your plot. Do not overwrite your original data.

```{r}

# Add your code here

```

## 3. If needed, use `select()`, `group_by()`, and `summarize()` to create a summary dataset for your visualization.

```{r}

# Add your code here

```


## 4. Use `ggplot2` to create a visualization of your data that addresses the question you identified above. Your plot should:  

  - use an appropriate variable on the x-axis
  - use an appropriate variable on the y-axis
  - include at least one geom layer
  - use color or grouping if it helps make the pattern clearer
  - include a title, subtitle, axis labels, and legend label when appropriate

```{r}

# Add your code here

```
