---
title: "Lab 03 - Degree Centrality"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

# Do not adjust these settings
# as they are here to set up 
# a few things for you.

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6)

```

<br>

```{r load_packages}

# make sure you have installed the SNACpack package
# devtools::install_github( "jacobtnyoung/SNACpack" )
# be sure to review the Getting Started with R chapter in
# the Social Network Analysis for Crime Analysts using R textbook
# if you have questions about how to install SNACpack
#
# Note that you cannot knit with an install.packages() function
# called in your RMD file. You only have to install packages
# once, then they are loaded using the library() function.

# load the packages you need
library( SNACpack )
library( network )
library( sna )

```

<br>

# Getting Started

## Setup

```{r clear_workspace}

# clear the workspace
rm( list = ls() )

```

<br>

## Part I: Working with an Undirected Network

<br>

### 1. Plot the network using this command: `gplot( london_gang_net )`.

```{r}

# Your code here

```

<br>

### 2. Describe how you would modify this plot to improve interpretability for a lay audience. Suggest at least <u>three changes</u> and explain why they matter. Be sure to pay attention to the type of network it is and how best to represent the information in the network.  

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 3. Implement the proposed changes using `gplot()` and describe how each change affects how we see the network.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 4. Without running any code, describe what kind of actor in this network is likely to have a *high* degree centrality score. 

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 5. Calculate the degree centrality score for each actor. 

```{r}

# Your code here

```

<br>

### 6. There are three actors with the highest degree centrality scores. Who are they and what does this say about their position in the network?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 7. What does it mean to "standardize" degree centrality? Why might we want to do this?
  
**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 8. Calculate the standardized degree centrality scores. Are the three actors with the highest degree centrality scores the same individuals with the highest *standardized* degree centrality scores? Why or why not?

```{r}

# standardized score is deg / g-1 where g is the number of nodes.

# add your code here

```
  
**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 9. Calculate the mean degree centrality score. Does the mean degree centrality score represent the "typical" actor in this network? Why or why not?

```{r}

# Your code here

```
  
**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 10. Graph centralization is often used to assess how "star-like" a network is. What would a completely centralized network look like? What would its degree centrality scores be? Describe it (or, of your feeling fiesty, create one and plot it!).

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 11. Calculate the graph centralization score for degree centrality. 
  
```{r}

# Your code here

```
  
<br>

### 12. Interpret the score. What does it tell us about the distribution of degree centrality scores in the network? Would you describe the network as centralized or decentralized? Justify your answer using both the score and the plot.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 13. Suppose you are an analyst in a gang unit in London. After analyzing this network, what might you do to address gang involvement? How is this informed by your knowledge of the network?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## Part II: Working with a Directed Network

<br>

### 1. Plot the network using this command: `gplot( cocaine_dealing_net )`.

```{r}

# Your code here

```

<br>

### 2. Describe how you would modify this plot to improve interpretability for a lay audience. Suggest at least <u>three changes</u> and explain why they matter. Be sure to pay attention to the type of network it is and how best to represent the information in the network.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 3. Implement the proposed changes using `gplot()` and describe how each change affects how we see the network.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 4. Without running any code, describe what kind of actor in this network is likely to have a *high* indegree centrality score and a *high* outdegree centrality score. How might that individual be different from someone who has a *high* indegree centrality score but a *low* outdegree centrality score? Which one is more prevalent in the network?  

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 5. Calculate the indegree and outdegree centrality scores for each actor. 


```{r}

# Your code here

```

<br>

### 6. There is one actor with the highest outdegree centrality score. What does this say about their position in the network? What are they doing that makes them have such a high score? (*hint*: think about what the ties measure).

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 7. Calculate the standardized indegree and outdegree centrality scores. 

```{r}

# Your code here

```

<br>

### 8. Calculate the mean indegree and outdegree centrality scores. Interpret each mean centrality score and compare them. Is one larger than the other? What does this mean?

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 9. Calculate the graph centralization score for indegree and outdegree centrality. 

```{r}

# Your code here

```

<br>

### 10. Interpret each score. What does it tell us about the distributions of indegree and outdegree centrality scores in the network? For indegree and outdegree, would you describe the network as centralized or decentralized? Justify your answer using both the score and the plot. Are the scores the same? What does this mean?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 11. Suppose you are an analyst in a drug unit examining cocaine trafficking. After analyzing this network, what might you do to address curb cocaine trafficking? How is this informed by your knowledge of the network?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## Part III: Your Networks

*Add your code setup here according to the instructions above, then DELETE THESE STATEMENT*

<br>
<br>

----

<br>
<br>


## *Challenge Question*

Feeling bold? Try this challenge question. It is not required, but try taking a crack at it.

*  Here is an edgelist for a small 6-node directed network:  
    * A → B  
    * A → C  
    * B → C  
    * C → A  
    * D → C  
    * E → F  

**Task:** Without using a computer, sketch this network. 

Then, answer:

  - Which node(s) have the highest indegree?  
  - Which node is most central by outdegree?  


**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  
