---
title: "Lab 04 - CLoseness/Betweenness Centrality"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

# Do not adjust these settings
# as they are here to set up 
# a few things for you.

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

```{r load_packages}

# load the packages you need
library( SNACpack )
library( network )
library( sna )

```

<br>

# Getting Started

## Setup

```{r clear_workspace}

# clear the workspace
rm( list = ls() )

```

<br>

## Part I: Working with an Undirected Network

<br>

### 1. Without running any code, describe what kind of actor in this network is likely to have a *high* **closeness** centrality score. Now, describe what kind of actor in this network is likely to have a *low* **betweenness** centrality score.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 2. Calculate the degree, closeness, and betweenness centrality scores for each actor. Feel free to use the `degree()`, `closeness()`, and `betweenness()` functions from the `sna` package.

```{r}

# Your code here

```

<br>

### 3. There are two actors with the highest closeness centrality scores. Who are they and what does this say about their position in the network? There is one actor with the highest betweenness centrality scores. Who are they and what does this say about their position in the network?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 4. Is the one actor with the highest betweenness centrality score one of the actors with the highest closeness centrality scores? What does this tell us about betweenness and closeness centrality for this network? Is the one actor with the highest betweenness centrality score one of the actors with the highest degree centrality scores? What does this tell us about betweenness and degree centrality for this network?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 5. What does it mean to "standardize" closeness centrality? What about betweenness centrality? Why might we want to do this?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 6. Calculate and report the standardized degree, closeness, and betweenness centrality scores for each actor.

```{r}

# Your code here

```

<br>

### 7. Are the actors with the highest closeness centrality scores the same individuals with the highest *standardized* closeness centrality scores? Why or why not?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 8. Are the actors with the highest betweenness centrality scores the same individuals with the highest *standardized* betweenness centrality scores? Why or why not?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 9. Calculate the mean closeness centrality score and the mean betweenness centrality score. Interpret each mean centrality score. What do these means suggest about the structure of the network?

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 10. Graph centralization is often used to assess how "star-like" a network is. If we were examining closeness centrality, what would a completely centralized network look like? What would its closeness centrality scores be? Describe it (or, if you are feeling spunky, create one and plot it!).

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 11. If we were examining betweenness centrality, what would a completely centralized network look like? What would its betweenness centrality scores be? Describe it (or, if your still feeling spunky, create one and plot it!).

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 12. Calculate the graph centralization scores for degree, closeness, and betweenness centrality.  

```{r}

# Your code here

```

<br>

### 13. Interpret each graph centralization score. What does each tell us about the distribution of centrality scores in the network? Would you describe the network as centralized or decentralized? Justify your answer using the score.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 14. Plot the network three times using the `gplot()` function where each plot sizes the nodes by the centrality score (i.e. a plot for degree centrality, a plot for closeness centrality, and a plot for betweenness centrality). Examine the plots. Describe how the differences you noted between the centrality measures above are seen in the plots.

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 15. Suppose you are an analyst in a gang unit in London. After analyzing this network, what might you do to address gang involvement? How is this informed by your knowledge of the network? Is your answer different based on whether you are examining degree centrality, closeness centrality, or betweenness centrality?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## Part II: Working with a Directed Network

<br>

### 1. Calculate the indegree, outdegree closeness, and betweenness centrality scores for each actor. Feel free to use the `degree()`, `closeness()`, and `betweenness()` functions from the `sna` package.

```{r}

# Your code here

```

<br>

### 2. Does the node with the highest closeness centrality score also have the highest betweenness centrality score? Explain your answer beyond "yes" or "no".

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 3. Look at the node with the highest indegree centrality score. Do they have the highest closeness centrality score? Betweenness centrality score? What does this say about their position in the network?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 4. Now, look at the node with the highest outdegree centrality score. Do they have the highest closeness centrality score? Betweenness centrality score? What does this say about their position in the network?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 5. Calculate and report the standardized indegree, outdegree, closeness, and betweenness centrality scores for each actor.

```{r}

# Your code here

```

<br>

### 6. Calculate and report the mean (unstandardized) indegree, outdegree, closeness, and betweenness centrality score. Interpret each mean centrality score. What do these means suggest about the structure of the network?

```{r}

# Your code here

```

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 7. Calculate the graph centralization scores for indegree, outdegree, closeness, and betweenness centrality.

```{r}

# Your code here

```

<br>

### 8. Interpret each score. What does it tell us about the distributions of each type of centrality score in the network? For each measure of centrality, would you describe the network as centralized or decentralized?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 9. Plot the network four times using the `gplot()` function where each plot sizes the nodes by the centrality score (i.e. a plot for indegree centrality, a plot for outdegree centrality, a plot for closeness centrality, and a plot for betweenness centrality).

```{r}

# Your code here

```

<br>

### 10. Examine the plots. Describe how the differences you noted between the centrality measures above are seen in the plots. If you can't see them, think about how to modify the plot to better visualize these differences.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 11. Suppose you are an analyst in a drug unit examining cocaine trafficking. After analyzing this network, what might you do to address curb cocaine trafficking? How is this informed by your knowledge of the network? Is your answer different based on whether you are examining indegree centrality, outdegree centrality, closeness centrality, or betweenness centrality?

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*

<br>

### 12. Now, suppose your colleague said "hey, this is a directed graph, but you can calculate the closeness and betweenness assuming it is an undirected graph you know?" Well, of course you know that, but should you? Think about what the directionality in the graph represents and then explain your answer.

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*



<br>
<br>

----

<br>
<br>

## Part III: Your Networks

*Add your code setup here according to the instructions above, then DELETE THESE STATEMENT*

<br>
<br>

----

<br>
<br>


## *Challenge Question*

Feeling bold? Try this challenge question. It is not required, but try taking a crack at it.

*  Here is an edgelist for a small 6-node directed network:  
    * A → B  
    * A → C  
    * B → C  
    * C → A  
    * D → C  
    * E → F  

**Task:** Without using a computer, sketch this network. 

Then, answer:

  - Which node(s) have the highest closeness centrality?  
  - Which node(s) have the highest betweenness centrality?  

**Response**: 
*ADD YOUR RESPONSE HERE (THEN DELETE THIS STATEMENT)*
  
<br>
<br>

----

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

  * Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
  * Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  
